Clustering based on probability distributions with application on residential customers

### 2021-10-20 main.Rmd
characterisation clusters updated

In the analysis of energy smart metre data, K-Means or hierarchical clustering are often employed. These are simple and effective techniques that work well in a range of scenarios. For clustering, both employ a distance measure, and the distance measure chosen has a major influence on the structure of the clusters. We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference). Individual entities with the highest similarity computed using the desired distance metrics are sequentially merged using agglomerative algorithms. We can possibly employ any clustering method that supports the given distance metric as input.

@@

Depending on the distance measure utilized for the study, the cluster characterization technique will differ. Clusters that utilise intra-category distances are characterised using multi-dimensional scaling and parallel coordinate displays. For inter-category distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met. This technique may potentially make advantage of multi-dimensional scaling.

\noindent Multidimensional scaling (MDS) (XXX reference) refers to a family of methods that analyse a matrix of distances or dissimilarities to provide a representation of the data points in a reduced-dimension space. There are many kinds of MDS, but they all solve the same fundamental issue: Given a $n*n$ matrix of dissimilarities and a distance measure, identify a configuration of $n$ points $x_1, x_2, \dots, x_n$ in the reduced dimension space $R^q$ ($q<p$) where the distance between the points is near to the dissimilarity between the points. All techniques must determine the coordinates of the points as well as the space dimension, $q$. Metric and nonmetric MDS are the two main kinds of MDS. Metric MDS methods presume a functional connection between the interpoint distances and the supplied dissimilarities and assume that the data are quantitative. We use metric MDS.

\noindent Parallel coordinate plots (XXX reference) Parallel coordinates have been extensively used to display high-dimensional and multivariate data, allowing for the detection of patterns within the data via visual grouping.

@@

All the methods were fitted to each data designs and results are reported through confusion matrices. With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs. For $mean_diff=1$, the performances are pretty bad for js-robust methods and wpd method for lower nT. Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example. It is likely that $nT$ would be at least $1000$ with half-hourly data, even if data is only available just for a month. The performance is promising except when the number of observations for a customer is really small. For smaller difference between categories, it is expected that method js-nqt would preform better than the other two.

>>>>

In the analysis of energy smart meter data, K-Means or hierarchical clustering are often employed. These are simple and effective techniques that work well in a range of scenarios. For clustering, both employ a distance measure, and the distance measure chosen has a major influence on the structure of the clusters. We employ agglomerative hierarchical clustering in conjunction with Ward's criteria (XXX reference). The pair of clusters with minimum between-cluster distance are are sequentially merged in this using this agglomerative algorithms. A good comprehensive list of algorithms can be found in @.Xu2015-ja. We can possibly employ any clustering method that supports the given distance metric as input.

@@

Characterization of clusters both statistically and qualitatively is an important stage of a cluster analysis. A potential way is to look at the findings from all the groups in graphs, and enhance our qualitative descriptions of the groupings. @Cook2007-qe provides several ways to characterize clusters.

(a) Parallel coordinate plot: Parallel coordinate plot (Inselberg 1985, Wegman 1990) are widely used to display high-dimensional and multivariate data, allowing visual grouping to detect patterns. In a Parallel Coordinates Plot, each variable has its own axis, which are all parallel. Each axis is connected by a series of lines. That is, each line is made up of connected points on each axis. The order of the axes can affect how the reader interprets the data. This is because adjacent variables are more easily perceived than non-adjacent variables and rearranging the axes can reveal patterns or correlations between variables. Scattered plots of the p variables are arranged in a scatterplot matrix. It's a neat way to show multiple relationships at once, and it allows us to compare all the plots at once.

(b) Scatterplot matrix: The scatter plot matrix (draftsman's plot) is a matrix that comprises pairwise scatter plots of the p variables.Pairwise scatter plots are excellent for determining relationships between variables and determining which factors have contributed the most to clustering.

(c) Plotting cluster statistics: For larger problems, parallel coordinate plots may become cluttered and difficult to read, therefore we may opt to display cluster statistics instead. (@Dasu_undated-gw)

(c) MDS, PCA and t-SNE: While all of the techniques examine a matrix of distances or dissimilarities to give a representation of the data points in a reduced-dimension space, their goals are not the same. The principal component analysis (Johnson & Wichern 2002) attempts to retain data variance. Multidimensional scaling (@borg2005modern) seeks to maintain the distances between pairs of data points, with an emphasis on pairings of distant points in the original space. t-SNE, on the other hand, is concerned with preserving neighbourhood data points. Close data points in high-dimensional space will be condensed in the t-SNE embeddings.

(d) Tour: A tour () is a collection of interpolated linear projections of multivariate data into a lower-dimensional space. The sequence is seen as a dynamic visualisation, enabling the viewer to observe the shadows cast by the high-dimensional data in a lower-dimensional view.
<!-- and also provides examples on how one might spin and brush using a tour, to build up a cluster analysis manually. -->

Depending on the distance measure utilized for the study, the cluster characterization technique will differ. Clusters that utilize wpd-based distances are characterised using multi-dimensional scaling and parallel coordinate displays. For JS-based distances, the distribution across major granularities may be presented to ensure that the goal of similar shapes within clusters and distinct shapes across clusters is met. This technique may potentially make advantage of multi-dimensional scaling.

@@

All the methods were fitted to each data designs and results are reported through confusion matrices. With increasing difference between categories, it gets easier for the methods to correctly distinguish the designs. For $mean_diff=1$, the performances are pretty bad for js-robust methods and wpd method for lower nT. Although, with the kind of residential load datasets, a full year of load is the minimal requirement to capture expected variations in winter and summer profiles, for example. It is likely that $nT$ would be at least $1000$ with half-hourly data, even if data is only available just for a month. The performance is promising except when the number of observations for a customer is really small. For smaller difference between categories, it is expected that method js-nqt would preform better than the other two.
<!-- A confusion table can come alive with linked brushing, so that mismatches and agreements between methods can be explored. -->
